---
title: "Introduction to demingx"
author: "Jan Gleixner"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to demingx}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#"
)
```

## Overview

The `demingx` package implements a heteroscedastic slope-only errors-in-variables (EIV) model.
It generalizes classical Deming regression by accounting for known per-observation uncertainties in both x and y,
as well as estimating an intrinsic scatter term \( \sigma_\eta^2 \).

## Simulation example

We simulate and fit data with known heteroscedastic errors:

```{r}
library(demingx)

set.seed(42)
d <- simulate_demingx(n = 50, beta = 1.5, sigma_eta = 0.2)
fit <- fit_demingx(d)
```

## Results

```{r}
fit
```

## Log-likelihood surface (optional)

```{r, fig.height = 4}
ll <- function(beta, log_sigma2_eta) {
  -demingx_loglik(c(beta, log_sigma2_eta), d)
}

beta_seq <- seq(1.2, 1.8, length.out = 100)
sigma_seq <- seq(log(0.01), log(0.3), length.out = 100)

grid <- expand.grid(beta = beta_seq, log_sigma2_eta = sigma_seq)
grid$loglik <- mapply(ll, grid$beta, grid$log_sigma2_eta)

library(ggplot2)
ggplot(grid, aes(beta, exp(log_sigma2_eta), z = loglik)) +
  geom_contour_filled() +
  scale_y_log10() +
  labs(y = expression(sigma[eta]), x = expression(beta))
```
